---
title: "Coronavirus disease (COVID-19) Pandemic"
author: "Kevin Bitinsky"
date: "March 25, 2020"
output: ioslides_presentation
---


## What is this Presentation?
This is an assignment for the Coursera / Johns Hopkins Data Science course, __Developing Data Prodcuts__. It is presentation using R Markdown that features a plot created with Plotly. 

To demonstrate the usage of R Markdown and Plotly, I have chosen to demonstrate how the COVID-19 pandemic has affected the world by showing the current infection statitstics on a Plotly chloropleth map.

## What is COVID-19?

According to the [World Health Organization](https://www.who.int/), 
>"COVID-19 is the infectious disease caused by the most recently discovered coronavirus. This new virus and disease were unknown before the outbreak began in Wuhan, China, in December 2019."

The data is courtesy of the [European Centre for Disease Prevention and Control](https://www.ecdc.europa.eu/)

```{r data_retrieval, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
# Data, and the following code snippet courtesy of :https://www.ecdc.europa.eu/
library(readxl)
library(httr)
url <- paste("https://www.ecdc.europa.eu/sites/default/files/documents/COVID-19-geographic-disbtribution-worldwide-",
             format(Sys.time(), "%Y-%m-%d"), 
             ".xlsx", sep = "")
GET(url, authenticate(":", ":", type="ntlm"), write_disk(tf <- tempfile(fileext = ".xlsx")))
data <- read_excel(tf)
##### End of code snippet
```

<!-- Manipulate the data to summarize Cases by Country. -->
<!-- I needed to groub_by both the *Country* name and the *GeoId* or else I seemed to be missing some, but not all, multiple word countries such as USA and UK (but not others such as UAE) when charting it in Plotly.  -->
<!-- * I could have updated the data for the USA and UK specifically, but could not be sure that I would not be missing some other countries. -->
```{r data_manipulation, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
library(dplyr)

colnames(data)[7] <- "Country"
df <- data %>% 
  group_by(Country, GeoId) %>%
  summarize(Cases = sum(Cases))
```


<!-- Added two traces, one for Country and one for GeoId -->
## The Plotly Chart
```{r plotly_creation, echo=FALSE, message=FALSE, warning=FALSE}
# Followed the model as demonstrated in the plotly documentation at: https://plotly.com/r/choropleth-maps/
library(plotly)
# dark grey boundaries
l <- list(color = toRGB("darkgrey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = TRUE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

fig <- plot_geo(df, locationmode = 'country names')
fig <- fig %>% add_trace(
  z = ~Cases, color = ~Cases, colors = 'Reds',
  text = ~Country, locations = ~Country, marker = list(line = l), showscale=FALSE
)  %>% add_trace(
  z = ~Cases, color = ~Cases, colors = 'Reds',
  text = ~Country, locations = ~GeoId, marker = list(line = l), showscale=FALSE
) 
  
fig <- fig %>% colorbar(title = "# of Cases")
fig <- fig %>% layout(
  title = paste('Cases of COVID-19', Sys.Date()),
  geo = g, 
  autosize = FALSE, #width = 900, height = 900,
  showlegend = TRUE,
  margin = list(
      r = 10, 
      t = 25, 
      b = 10, 
      l = 10
    )
)

fig
```


## References

- [The Plolty Website](https://plot.ly/)
- [The Plotly World Chloropleth Map](https://plotly.com/r/choropleth-maps/)